---
title: "Mousetracking Analysis"
author: "Laura Bock Paulsen"
date: "9/11/2022"
output: html_document
---

For the analysis of your very own mouse-tracking data, we will be using a package called MouseTrap. In order to get acquainted with the package, we will play around with data from a single participant conduction the circle-square experiment. 

The documentation for the package can be found here: <https://www.rdocumentation.org/packages/mousetrap/versions/3.1.5/topics/mousetrap>.

Use this page to solve the following steps by finding and applying appropriate mousetrap functions. Also, writing ?function_name() in the console is great way of reading about the function in question.

Call me for help whenever you get stuck (or take a quick look in the solutions file) :-)


### Install packages and load in the data
```{r}
pacman::p_load(tidyverse, mousetrap, gridExtra, readbulk, arsenal)
```

```{r}
df_mouse <- read_opensesame(directory = "subjects")

#listm <- list.files(path = "Mousetracking/Subjects" , pattern = "subject")
#listm = list.files(pattern="*.csv")
#myfiles = lapply(listm, read.delim)

#df <- do.call(rbind.data.frame, myfiles)

```

```{r}
df_mouse$timestamps_mouse_1 <- NULL
df_mouse$xpos_mouse_1 <- NULL
df_mouse$ypos_mouse_1 <- NULL
```


### Turn the data into a mousetrap object
Take a moment to look at the structure that it created. What are the elements, what information does it provide?
```{r}

mt_obj <- mt_import_mousetrap(df_mouse)
```

### Make a quick plot using the mt_plot() function
Try to understand what the lines mean and how to “read” the graph. What seems weird about it when you think about how the experiment task looked like?
```{r}
group_colors <- c(blue_button = "#1458e8", green_button = "#249a07", orange_button = "#ec8019", violet_button = "#a919ec")

plot2 <- mt_plot(mt_obj, use = "trajectories", color = "response_mouse")

trajectory_plot <- plot2 + ggplot2::scale_colour_manual(values=c(blue_button = "#1458e8", green_button = "#249a07", orange_button = "#ec8019", violet_button = "#a919ec")) + ggtitle("Mousetracking Trajectory") + theme(plot.background = element_rect(fill = "grey80", colour = NA))
trajectory_plot
ggsave("traject.jpg", trajectory_plot)

plot3 <- mt_plot_aggregate(mt_obj)
plot3
```


```{r}
group_colors
```

```{r}
word_colors <- 

plot45 <- mt_plot(mt_obj, use = "trajectories", color = "Word", size = 0.5)

trajectory_plot2 <- plot45 + ggplot2::scale_colour_manual(values=c(Abortion = "#a83221", Compiling... = "#a83221", Death = "#a83221", Guilt =  "#a83221", Pain = "#a83221", Slaughter = "#a83221", Methods = "#a83221", Portfolio = "#a83221", Bagel = "#249a07", Chair = "#249a07", Dog = "#249a07", Finger = "#249a07",  Glasses = "#249a07", Mountain = "#249a07", Shirt = "#249a07", Water = "#249a07")) + ggtitle("Mousetracking Trajectory") + theme(plot.background = element_rect(fill = "grey80", colour = NA))
trajectory_plot2
ggsave("traject2.jpg", trajectory_plot2)

 #249a07 
 #a83221 

```


### Find a function that removes the initial phase without mouse-movement
Find a function to remove this “line” and plot the timestamps by xpos again. 
```{r}


m <- mt_exclude_initiation(mt_obj)
```



**Thought experiment:** If you now plotted x and y coordinates again, what would have changed in the plot? Think first, then try it! 
```{r}
plot_xy <- mt_plot(m, use = "trajectories", x = "timestamps",y ="xpos", color = "Word")
#plot_time
plot_xy
```

### Time-normalize the data
Apply the function “mt_time_normalize”
```{r}
#Re-represent each trajectory spatially using a constant number of points so that adjacent points on the trajectory become equidistant to each other
time_norm <- mt_time_normalize(m, save_as = "tn_trajectories")

```

Now look at your data variable (the mt object) where a new matrix appeared. What do you notice? What does the function do? After you have thought about it yourself, check the documentation!

### Plot the normalized trajectories
Find out how to plot the normalized trajectories instead of the raw data.
```{r}

plot_tn <- mt_plot(time_norm, use = "tn_trajectories", x = "timestamps",y ="xpos", color = "Word")
plot_tn
```

### Play around with the mt_time_normalize() function
Take a moment to play around with different numbers of steps in “mt_time_normalize” and see how that changes the shape of the trajectories, esp. when using very small numbers. Discuss what the decision about normalizing implies for the data analysis and interpretation. In the end, set the steps back to the default of 101.
```{r}
# nsteps an integer specifying the number of equally sized time steps
tn2 <- mt_time_normalize(remove_init, nsteps = 101)

plot_tn2 <- mt_plot(tn2, use = "tn_trajectories", x = "timestamps",y ="xpos", color = "trial_type")
plot_tn2
```

### Aggregated plots
Now we want to visualize our “findings”. Find a function that will plot averages of all the “similar” movements and all the “dissimilar” movements.
Think: Which trajectories do we need to use, the original or the time normalized? Why? Try plotting both to see whether you were right.
```{r}
agg <- mt_aggregate(m, use = "trajectories")
agg_tn <- mt_aggregate(time_norm, use = "trajectories" )

plot_agg <- mt_plot_aggregate(m, color = "Word")
plot_agg

plot_agg_tn <- mt_plot_aggregate(time_norm, use = "tn_trajectories", color = "Word") + labs(title = "time-normalized")
plot_agg_tn


```

### Apply the function mt_measures()
Apply the function “mt_measures” and look at the outcome in your data variable. 
```{r}

meas <- mt_measures(time_norm)
meas
```

### Aggregating measures
Now find a function that helps you aggregate some measures of your pleasing over the two trial_types.
```{r}

meas_agg <- mt_aggregate(meas, 
                         use = "measures", 
                         use_variables = c("RT", "idle_time", "MAD_time"),
                         use2_variables = "Word"
                         )
meas_agg

```



```{r}
vec2 <- c("con", "con", "con", "met","con", "con", "con", "con","con","met","con","con","met", "con","con","con")

data3 <- meas_agg

data3$valence <- vec2    

```



```{r}
vec1 <- c("neg", "con", "con", "met","neg", "con", "con", "con","neg","met","con","neg","met", "con","neg","con")

data2 <- meas_agg

data2$valence <- vec1    

```


```{r}
vec <- c("neg", "con", "con", "neg","neg", "con", "con", "con","neg","neg","con","neg","neg", "con","neg","con")

data1 <- meas_agg

data1$valence <- vec    

```


```{r}
one_way <- aov(MAD_time ~ valence, data = data2)

one_way2 <- aov(RT ~ valence, data = data2)

```



```{r}
summary(one_way)
summary(one_way2)


```


```{r}
library(dplyr)
group_by(data1, valence) %>%
  summarise(
    count = n(),
    mean = mean(RT, na.rm = TRUE),
    sd = sd(RT, na.rm = TRUE)
  )
```


```{r}
library(dplyr)
group_by(data1, valence) %>%
  summarise(
    count = n(),
    mean = mean(MAD_time, na.rm = TRUE),
    sd = sd(MAD_time, na.rm = TRUE)
  )


```


```{r}
library(dplyr)
group_by(data2, valence) %>%
  summarise(
    count = n(),
    mean = mean(RT, na.rm = TRUE),
    sd = sd(RT, na.rm = TRUE),
  )


```


```{r}
library(dplyr)
group_by(data2, valence) %>%
  summarise(
    count = n(),
    mean = mean(MAD_time, na.rm = TRUE),
    sd = sd(MAD_time, na.rm = TRUE),
  )

```






```{r}
# Shapiro-Wilk normality test for Men's weights
with(data1, shapiro.test(MAD_time[valence == "neg"]))# p = 0.1
# Shapiro-Wilk normality test for Women's weights
with(data1, shapiro.test(MAD_time[valence == "con"])) # p = 0.6

```




```{r}
res.ftest <- var.test(RT ~ valence, data = data1)
res.ftest


```


```{r}

res <- wilcox.test(RT ~ valence, data = data1,
                   exact = FALSE)
res
```

```{r}
res1 <- wilcox.test(MAD_time ~ valence, data = data1,
                   exact = FALSE)
res1

```


```{r}
res3 <- wilcox.test(RT ~ valence, data = data3,
                   exact = FALSE)
res3

```



```{r}


sum_neg <- sum(meas_agg[which(meas_agg$Word == 'Abortion' | meas_agg$Word =='Death' | meas_agg$Word =='Compiling...' | meas_agg$Word =='Guilt' | meas_agg$Word =='Methods' | meas_agg$Word =='Pain' | meas_agg$Word =='Portfolio' | meas_agg$Word =='Slaughter'), 2])

sum_neg/8

sum_pos <- sum(meas_agg[which(meas_agg$Word == 'Bagel' | meas_agg$Word =='Chair' | meas_agg$Word =='Dog' | meas_agg$Word =='Finger' | meas_agg$Word =='Glasses' | meas_agg$Word =='Mountain' | meas_agg$Word =='Shirt' | meas_agg$Word =='Water'), 2])

sum_pos/8


```



### Interpret the results
Take a moment to think what these “results” could mean if this was a proper experiment which was not run by your instructor. How would you interpret this? Does this match your hypothesis? Especially look at the MAD (maximum absolute deviation from the direct path), the AUC (area under the curve) and the reaction time. 

It matches the hypotheses that the reaction times will generally be longer when the trial type is similar, meaning that similar stimuli crave more time to properly identify. 
On a larger scale, the results from our only participant would probably be corroborated within a bigger sample size. 


## Additional exercises

### Removing incorrect trials
Normally, you'd want to remove incorrect trials. As the mouse_trap object does not function with tidyverse, figure out how to remove potential incorrect trials from your mousetrap object. 
```{r}

```

### Discussion of mt_align
Would the function ‘mt_align’ be useful for this data? Why or why not?
```{r}

```

